---
title: "State Covid Testing Per Capita | Data Analysis"
author: "Brenda Wintrode"
date: "05/17/2020"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, paged.print=TRUE)
```

## Libraries and Packages

```{r}
# Load Tidyverse, janitor and USAboundaries, tidycensus
#install.packages("ggplot2")
library(tidycensus)
library(tidyverse)
library(janitor)
library(USAboundaries)
library(ggplot2)
library(scales)
library(mapview)
library(ggthemes)

```



```{r}
#USAboundaries changed the state names to two-letter abbreviations. I needed to glimpse it here to see what the variable names were.
glimpse(state_codes)
```

## Cleaned Data

```{r}
#Scraped the state by state data from The Covid Tracking Project.
covid_data <- read_csv("https://covidtracking.com/api/v1/states/current.csv")
```



```{r}
#Made all column names lower case and separated by underscores.
#Selected just the total test results and state name columns.
covid_data_working <- covid_data %>%
  clean_names() %>%
  select(state, total_test_results)
  glimpse(covid_data_working)
```


```{r}
#Loaded the census data API and the state population variable code.
census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")
#acs_variables <- load_variables(2017, "acs5" )

state_population <- get_acs(geography = "state", 
              variables="B01001_001")
```

```{r}
#Cleaned the state population data. To get mutate to work, I had to take our Puerto Rico because USAboundaries wouldn't work on Puerto Rico.
#Filtered out the states whose data grade was a C or below. Some B-grade states were filtered out because total test results were not per person, but instead per specimen or did not include private lab totals. States with B grades that were filtered out were Indiana, Iowa, Maine, Massachusetts, California.
state_population_working <- state_population %>%
 clean_names() %>%
  filter(name != "Puerto Rico") %>% 
  mutate(state = state_codes$state_abbr[state_codes$state_name==name]) %>%
 filter(!(state %in% c("ME","PA")))
  
``` 
## Inner Join: Tests Per Capita

*Joined the total tests and state populations on state abbreviation.
*Created a new column with a per state calculation of percent of poulation tested.

```{r}

tests_per_state_population <- covid_data_working %>%
  inner_join(state_population_working, by=c("state" = "state")) %>%
  mutate(percent_of_population_tested = (total_test_results/estimate)*100)
 
  
```

## Map Making
```{r}
#Pulling in state geometry like we did in Lab 07.

state_geodata <- get_acs(geography = "state",
              variables = "B01001_001", geometry = TRUE)
```
``` {r}
#Adjusting for Hawaii and Alaska

state_geodata_shifted <- get_acs(geography = "state",
              variables = "B01001_001", geometry = TRUE, shift_geo = TRUE)
              
```

```{r}

print(head(state_geodata_shifted))

```

```{r}
state_geodata_shifted %>%
  clean_names()
```

## Mapped Findings

``` {r}
testing_percentage_map <- state_geodata_shifted %>%
  inner_join(tests_per_state_population, by=c("GEOID" = "geoid"))

```

```{r}
testing_percentage_map %>%
  ggplot(aes(fill = percent_of_population_tested)) +
  geom_sf(lwd = 0) +
  theme_map() +
  labs(fill='Percentage',title="Percentage of COVID-19 Tests Administered by State Population", subtitle = "2017 population, U.S. Census", caption = "Source: U.S. Census ACS and The Covid Tracking Project") +
  theme(legend.position="right") +
  scale_fill_viridis_c(option = "magma",labels = comma)
  
# Turn off scientific notation if your legend looks weird
options(scipen=999)
```

## Hover Map
* According to The Covid Tracking Project data grades, Deleware, Pennsylvania and Maine were not tracking total testing

``` {r}
mapview(testing_percentage_map, zcol = "percent_of_population_tested", legend = FALSE, legend.opacity=.1)

```